Automatic Bicycle Balance Assistance Reduces Probability of Falling at Low Speeds When Subjected to Handlebar Perturbations\\Version 4.dev

### 2025-11-18 main.tex
Switch to bibtex instead of biblatex + biber.
bibtex instead of biblatex now working.
Added sentence explaining the exclusion of two participants' data.

\usepackage{amsmath}
\usepackage{booktabs}  % nice tables
\usepackage[margin=25mm]{geometry}
\usepackage[natbib=true,style=numeric,sorting=none,giveninits=true]{biblatex}
% NOTE : this file is automatically generated from Zotero, do not edit
% manually!
\addbibresource{references.bib}
\usepackage{siunitx}  % use for all units
\usepackage{subcaption}  % for subfigures
\usepackage[]{todonotes}  % use [disable] to hide notes

@@

Following the initial threshold determination where the participant experienced
a full range of perturbation magnitudes, we then choose perturbation forces
according to a random and adaptive staircase procedure applying perturbations
above and below the initial perturbation threshold, while allowing small
progression of the perturbation threshold to accommodate learning effects. We
did not try to eliminate learning effects due to the time per participant that
would be required, but factored this into the later analysis as an alternative.
Five possible perturbation forces are determined based on the initial
perturbation threshold estimate: the initial estimate itself, two perturbations
lower than the initial estimate, and two perturbations higher than the initial
estimate.  The five possible perturbations are separated by 10~\si{\newton}
steps. For example, if the initial estimate of the perturbation threshold of a
participant is 80~\si{\newton}, the five possible perturbations are 60, 70, 80,
90 and 100~\si{\newton}. Which one of these five perturbations is chosen first
is determined at random. If the perturbation results in a fall, the estimate of
the perturbation threshold is decreased by 10~\si{\newton}, and vice versa if
the perturbation did not result in a fall. Five new possible perturbation
forces are determined around the updated perturbation threshold, and a new
perturbation is selected at random. This process iterates until twenty
perturbations are applied. The goal of this adaptive staircase procedure is to
have participants fall for approximately 50\% of the time. After the first set
of 20 perturbations, we let the cyclist rest and then perform another 20
perturbations. We randomize whether the balance assist system is on or off
during the first or second set of 20 perturbations for each participant and the
participants were not told the order.

@@

Before fitting the model, we scale each independent variable \(x_{ij}^k\) such
that they have a mean of zero and a standard deviation of one by cluster-mean
centering with clusters being an individual participant, as recommended by
\cite{Enders2007}. The clusters are chosen as all data from an individual
participant because we are only interested in the association between the state
of the balance assist system and the outcome of the perturbation for that
participant, not how participants perform relative to each other. We expected
there to be variation between participants in how well they are able to resist
the perturbation; however, this was not true. If a random effect is included in
the model for each participant, the intraclass correlation (ICC) is less that
3\% for both the 6~\si{\kph} and 10~\si{\kph} datasets, so we utilize a simple
single-level logistic regression model without the random effect instead of a
multilevel model. This tracks with the independent data collected in
\cite{Haitjema2023a} which also had negligible participant variation. This left
us with angular impulse, perturbation order, balance assist state, and roll \&
steer angles at the time of perturbation as independent variables. We also
include interaction effects between the balance assist state and the other four
variables. We divide the analysis into two separate model evaluations, one for
the 6~\si{\kph}, \(\kappa=5.2\) trials and one for the 10~\si{\kph},
\(\kappa=3.9\) trials and we evaluate our hypothesis for each set of data.
Statistics were computed with R version 4.1.3~\cite{RCoreTeam2021} the lme4
version 1.1~\cite{Bates2015} package.

@@

\printbibliography

>>>>

\usepackage{amsmath}
\usepackage{booktabs}  % nice tables
\usepackage[margin=25mm]{geometry}
\usepackage{siunitx}  % use for all units
\usepackage{subcaption}  % for subfigures
\usepackage[]{todonotes}  % use [disable] to hide notes
\usepackage[hidelinks]{hyperref}

@@

Following the initial threshold determination where the participant experienced
a range of perturbation magnitudes, we then choose perturbation forces
according to a random and adaptive staircase procedure applying perturbations
above and below the initial perturbation threshold, while allowing small
progression of the perturbation threshold to accommodate learning effects. We
did not try to eliminate learning effects due to the time per participant that
would be required, but factored this into the later statistical analysis.  Five
possible perturbation forces are determined based on the initial perturbation
threshold estimate: the initial estimate itself, two perturbations lower than
the initial estimate, and two perturbations higher than the initial estimate.
The five possible perturbations are separated by 10~\si{\newton} steps. For
example, if the initial estimate of the perturbation threshold of a participant
is 80~\si{\newton}, the five possible perturbations are 60, 70, 80, 90 and
100~\si{\newton}. Which one of these five perturbations is chosen first is
determined at random. If the perturbation results in a fall, the estimate of
the perturbation threshold is decreased by 10~\si{\newton}, and vice versa if
the perturbation did not result in a fall. Five new possible perturbation
forces are determined around the updated perturbation threshold, and a new
perturbation is selected at random. This process iterates until twenty
perturbations are applied. The goal of this adaptive staircase procedure is to
have participants fall for approximately 50\% of the time. After the first set
of 20 perturbations, we let the cyclist rest and then perform another 20
perturbations. We randomize whether the balance assist system is on or off
during the first or second set of 20 perturbations for each participant and the
participants were not told the order.

@@

Due to data quality issues that prevented time synchronization of the
measurements, we excluded one participant from the 6~\si{\kps} dataset and one
from the 10~\si{\kps} dataset, leaving 10 and 14 included participants with 40
perturbations each. Before fitting the model, we scale each independent
variable \(x_{ij}^k\) such that they have a mean of zero and a standard
deviation of one by cluster-mean centering with clusters being an individual
participant, as recommended by \cite{Enders2007}. The clusters are chosen as
all data from an individual participant because we are only interested in the
association between the state of the balance assist system and the outcome of
the perturbation for that participant, not how participants perform relative to
each other. We expected there to be variation between participants in how well
they are able to resist the perturbation; however, this was not true. If a
random effect is included in the model for each participant, the intraclass
correlation (ICC) is less that 3\% for both the 6~\si{\kph} and 10~\si{\kph}
datasets, so we utilize a simple single-level logistic regression model without
the random effect instead of a multilevel model. This tracks with the
independent data collected in \cite{Haitjema2023a} which also had negligible
participant variation. This left us with angular impulse, perturbation order,
balance assist state, and roll \& steer angles at the time of perturbation as
independent variables. We also include interaction effects between the balance
assist state and the other four variables. We divide the analysis into two
separate model evaluations, one for the 6~\si{\kph}, \(\kappa=5.2\) trials and
one for the 10~\si{\kph}, \(\kappa=3.9\) trials and we evaluate our hypothesis
for each set of data.  Statistics were computed with R version
4.1.3~\cite{RCoreTeam2021} the lme4 version 1.1~\cite{Bates2015} package.

@@

\bibliographystyle{plain}
% NOTE : this file is automatically generated from Zotero, do not edit
% manually!
\bibliography{references}
